SHYFEM includes an as\-tro\-no\-mi\-c tidal model which can be
 activated by setting the parameter |rtide| equal 1 in the |para| 
 section.

 The model calculates equilibrium tidal potential ($\eta$) and load 
 tides ($\beta$) and uses these to force the free surface. 
 The term $\eta$ in the momentum equations is calculated as a sum
 of the tidal potential of each tidal constituents multiplied by the
 frequency-dependent elasticity factor. The factor $\beta$ accounts 
 for the effect of the load tides, assuming that loading tides are
 in-phase with the oceanic tide. $\beta$ is function of the water 
 depth as $\beta=ltidec*H$ with |ltidec| a calibration factor to be 
 set in the str |para| section.

 The model cosiders the following tidal costituents:
 \begin{itemize}
 \item Semidiurnal species:
    \begin{itemize}
    \item M2  semi-diurnal principal lunar
    \item S2  semi-diurnal principal solar
    \item N2  large elliptical tide of first-order to M2
    \item K2  semi-diurnal declination to M2
    \item NU2 large evection tide to M2
    \item MU2 large variation tide to M2
    \item L2  small elliptical tide of first-order to M2
    \item T2  large elliptical tide of first-order to S2
    \end{itemize}
 \item Diurnal species:
    \begin{itemize}
    \item K1  declination luni-solar
    \item O1  principal lunar
    \item P1  principal solar
    \item Q1  elliptical lunar
    \item J1  elliptical tide of first-order to K1
    \item OO1 evection tide to O1
    \item S1  radiational tide 
    \end{itemize}
 \item Long-Period species:
    \begin{itemize}
    \item MF  fortnightly lunar
    \item MM  monthly lunar
    \item MSM S0-semiannual solar
    \item MSF Evection tide to M0
    \item SSA semiannual solar
    \item SA  elliptical tide of first-order to S0
    \end{itemize}
 \end{itemize}

 SHYFEM also allows to perform the tidal analysis of water levels
 during the model runtime. The tidal analysis is actived by setting
 |itmtid| and |idttid|. |idttid| should be long enough to perform 
 a reliable analysis. The parameter |itmtid| can be used to start 
 the analysis after the simulation spin-up. The tidal analysis module
 write an output file .tide.shy containing amplitudes and phases
 of all tidal constituents over the computational domain (on the 
 nodes).
